\section{timeline}
In this section I show the specific timeframe for each
project along with the objectives to be addressed. The first two items
correspond to the work presented for my first comps. I included those
papers in my M.Sc. thesis (defended in January, 2014).


\begin{itemize}
  \item Understanding reverse time migration backscattering: noise or signal?
  
        Status: completed (Q3 2011-Q4 2013)
          
        Published: \cite{DiazRTM}

  \item Wavefield tomography using RTM backscattering (completed: \cite{diaz2015})

        Status: completed (Q3 2011-Q2 2014)
          
        Published:  \cite{diaz2015}


  \item {1- CIP wavefield tomography with illumination compensation (in progress)}
        
        Project steps: 
          \begin{itemize}
            \item Theory and testing with synthetic examples: completed.
            \item Application to real data: completed.
            \item Writing of the manuscript: ongoing (projected finish: December 10, 2015).
            \item To be submitted to Geophysics or Geophysical Prospecting.
          \end{itemize}

  \item {2- Seismic tomography using local correlation functions (in progress, \cite{DiazLcorr})}

        Project steps: 
          \begin{itemize}
            \item Theory and testing: completed.  
            \item Examples: ongoing (80\% completed) 
            \item To be submitted to Geophysics or Geophysical Prospecting.
          \end{itemize}

        
  \item {3- Extended imaging using Marchenko wavefields (in progress) }


        Project steps: 
          \begin{itemize}
            \item Theory and testing: ongoing:
            \begin{itemize}
                \item Test to background velocity errors :ongoing .
                \item Analysis of reflectivity dependence to scattering angle (AVA): ongoing.
            \end{itemize}  
            \item To be submitted to Geophysics or Geophysical Prospecting.
          \end{itemize}  

  \item {4- Wavefield tomography using multiple scattered waves (in progress) } 


        Project steps: 
          \begin{itemize}
            \item Theory and testing: ongoing:
            \begin{itemize}
                \item Sensitivity kernels based on multiple scattered waves: ongoing.
            \end{itemize}  
            \item To be submitted to Geophysics or Geophysical Prospecting.
          \end{itemize}  

\end{itemize}
\inputdir{.}
\plot{Gantt}{width=\textwidth}{ Gantt chart showing the progress of the
projects discussed in this thesis. The black line within each progress
box shows the completed percentage of the project.}
