%% start of file `template.tex'.
%% Copyright 2006-2012 Xavier Danaux (xdanaux@gmail.com).
%
% This work may be distributed and/or modified under the
% conditions of the LaTeX Project Public License version 1.3c,
% available at http://www.latex-project.org/lppl/.
\documentclass[11pt,letter]{moderncv}
\moderncvcolor{blue}   
\usepackage[utf8]{inputenc}
\usepackage[none]{hyphenat}
\usepackage{listings}
\newcommand{\mailt}[1]{\mail{^\textrm{\scriptsize#10}}}
\hyphenpenalty 5000
\exhyphenpenalty 500
\usepackage[margin=0.8in]{geometry}

\moderncvstyle{classic}                        % style options are 'casual' (default), 'classic', 'oldstyle' and 'banking'
%\nopagenumbers{}           
% the ConTeXt symbol
\def\ConTeXt{%
  C%
  \kern-.0333emo%
  \kern-.0333emn%
  \kern-.0667em\TeX%
  \kern-.0333emt}
\definecolor{web}{rgb}{0.2,0.2,0.2}
% adjust the page margins
\setlength{\hintscolumnwidth}{0.17\textwidth}
\setlength{\separatorcolumnwidth}{0.025\textwidth}

\renewcommand{\listitemsymbol}{--}

\firstname{Esteban F. Díaz Pantin}
\familyname{ }
\mobile{{720-343-92-32}}
\email{ediazpantin@gmail.com}
\newcommand{\up}[1]{\ensuremath{^\textrm{\scriptsize#1}}}

\begin{document}
\makecvtitle
\thispagestyle{empty}
\cvline{}{Exploration geophysicist focused in  seismic data processing, depth imaging, and wave equation velocity analysis.}
\section{{Education}}
\cventry{Aug. 2013--Dec 2016 }{\textbf{PhD in Geophysics}}{Colorado School of Mines}{Golden, CO, USA}{}{%
Thesis title: Extended imaging and tomography under two-way operators, GPA 3.78/4}
\cventry{Aug. 2011--Jan. 2014}{\textbf{M.Sc. in Geophysics}}{Colorado School of Mines}{Golden, CO, USA}{}{%
  Thesis title: RTM backscattering, noise or signal? GPA 3.76/4}
\cventry{Oct. 2002--Dec. 2007}{\textbf{Geophysical Engineering}}{Universidad Simón Bol\'{i}var}{Caracas, Venezuela}{}{GPA 3.94/5}

\section{{Experience}}

\cventry{Dec 2016-- }{\textbf{Geophysicist}}{Subsurface Technology, BP}{Houston, USA}{}{ Challenger Level J}

\cventry{May 2014-- Aug. 2014}{\textbf{Summer intern}}{Imagerie Sismique, Total}{}{Pau, France}{%
Constrained optimization applied to Full Waveform Inversion (FWI). His tasks involved the development, implementation,
and testing of the optimization approach in the production code.}

\cventry{ May 2013-- Aug. 2013}{\textbf{Summer intern}}{Subsurface Technology, BP}{}{Houston, USA}{%
Data preconditioning for full waveform inversion. His tasks included a 
successful test with a 3D field dataset.}

\cventry{ Aug. 2011-- Present}{\textbf{Research Assistant}}{}{Center for Wave Phenomena, Colorado School of Mines}{}{%
Wave equation migration velocity analysis using two-way operators. Esteban
 also maintains cluster and inversion tools within Madagascar users at CWP.}

\cventry{ Aug. 2009--Jul. 2011}{\textbf{Geophysicist}}{}{GeoImaging Soluções Tecnológicas LTDA, Rio de Janeiro, Brazil}{}{%
   Participated in several projects including Full Waveform Inversion R\&D, depth imaging,
   velocity model building under salt environments, and time processing.
}

\cventry{ Jan. 2009--Aug 2009}{\textbf{Geophysicist}}{}{Stratageo Soluções LTDA, Rio de Janeiro, Brazil}{}{%
   Time processing, depth imaging and velocity model building.  Seismic data from Brazil (onshore and offshore) and West Africa basins (offshore).
   Developed software to automatically process P1/90 navigation data. 
}

\cventry{ Mar. 2008--Dec. 2008}{\textbf{Junior Geophysicist}}{}{3DGeo South America, Buenos Aires, Argentina \& Rio de Janeiro, Brazil}{}{%
   Seismic data processing from Argentinian basins.
   Geometry QC and coordinates processing. 
   Near Surface model building.
}

\section{{Languages}}
\cvlanguage{Spanish}{Native}{}
\cvlanguage{English}{Full professional proficiency}{}
\cvlanguage{Portuguese}{Professional working proficiency}{}
\cvlanguage{French}{Elementary proficiency}{}
\section{{Computational Skills}}
\cvlistitem{Contributor to the Madagascar open-source project: www.ahay.org}
\cvlistitem{Programing: Fortran90, C, Python, Java, bash, Matlab, Javascript,among others.}
\cvlistitem{Developing experience within SEPlib, MinesJTK, and Madagascar environments.}

\section{{Scientific Interests}}
\cvlistdoubleitem{Migration Velocity Analysis (MVA)}{Inverse problems}
\cvlistdoubleitem{Imaging}{Noise reduction in seismic data}
% One page resume, uncomment for publications
\bibliographystyle{plain}
\bibliography{Esteban2}
\nocite{*}
\thispagestyle{empty}
\end{document}


%% end of file `template.tex'.
