\documentclass[10pt]{article}
\usepackage{amsmath, amsthm, amsbsy, rotating,float}
\usepackage{graphicx,algorithm,algorithmic}
\usepackage{setspace,enumerate}
\usepackage[margin=1.2in]{geometry}


\usepackage{graphicx,caption,subfig}

\floatstyle{ruled}
\newcommand{\norm}[1]{\left\lVert#1\right\rVert_p}
\newcommand{\normi}[1]{\left\lVert#1\right\rVert_\infty}
\newcommand{\normt}[1]{\left\lVert#1\right\rVert_2}
\doublespacing

\author{Esteban D\'{i}az}
\title{How can we increase the energy independence of the United States?}{}

\begin{document}
\date{February 20, 2014}
\maketitle
How much energy do we need?

Can we reduce our energy consumption?

Can we increase our energy production?

What energy alternatives should we look into?

Can we explore space for resources?

Can we be more energy responsible citizens?

What are other countries doing for energy independence?

Can we do the same? 

Is bio-fuels a plausible alternative?

Is it responsible to turn agriculture oriented farming to energy
oriented farming?

What about moral dilemmas like food starvation while we use our
fields to produce bio-fuels?

Should we open exploration on the west and the east coast? 

What are the implications?

Is our society ready for compromises? 

What are the energy needs for the next 40 years?

What can we do now to reduce them?

Is it fracking/unconventionals the best alternative?

Why? 

What can we do to inform the people about fracking technology? 

What about nuclear energy? 

Is there a way to make it right? 

What can we learn from other countries accidents? 

How can we include renewable energy in our day to day life?

Can we substitute conventional electricity generation with
house roof photovoltaic cells?

What about Geothermal? Can we make sustainable towns around
geothermal areas?

Can we separate fossil energy requirements from non-fossil requirements? 
(For instance, we could generate electricity with wind, water, etc; But not
flight planes with that)

Can we separate each sector requirements and set specific independence
goals?

What is the role of the State? What should be the role of the State?

If we let the development of new sources or initiatives in private hands,
 are we going to be independent?

\end{document}
